Dark matter halo shape at $z=0$ in the Auriga simulations: radial evolution and alignment with the stellar disk

### 2019-07-18 Document/paper.tex
new plot

\title[The shape of dark matter haloes at $z=0$ in Auriga]
{Dark matter halo shape at $z=0$ in the Auriga simulations: radial
  evolution and alignments with the stellar disk}
\author[Jesus Prada,  Jaime E. Forero-Romero, Volker Springel ]{
Jesus Prada,$^{1}$\thanks{E-mail: jd.prada1760@uniandes.edu.co}
Jaime E. Forero-Romero,$^{1}$
Volker Springel$^{2}$
\\
% List of institutions
$^{1}$Departamento de F\'sica, Universidad de los Andes, Cra. 1 No.
18A-10, Edificio Ip, Bogot\'a, Colombia.\\
$^{2}$Heidelberg Institute for Theoretical Studies,
Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany.\\
}

@@

% Enter the current year, for the copyright statements etc.
\pubyear{2018}

@@

% Abstract of the paper
\begin{abstract}
We present shape measurements of dark matter halos at redshift $z=0$
in a suite of 30 cosmological simulations from the Auriga project.
For all measurements we compare the results in full
magnetohydrodynamics against dark matter only physics.
We find a strong influence of baryons in making dark matter halos
rounder at all radii compared to its dark matter only counterparts.
This change in triaxility its more pronounced towards the inner
regions of the halo.
In simulations with baryons we measure a twist in the dark matter
shape as a function of radial distance.
We present this effect as a changing degree of halo shape alignment
with the stellar disk shape as a function of radius; the stronger
alignment is found at $0.25R_{200}\sim56$ kpc. 
This twisting in halo shape is almost absent from dark matter only
simulations. 
We show that a coherent halo shape with radius correlates with low
triaxility values.  
At distances close to the disk, rounder and coherent dark matter
distributions correlate with massive stellar disks and low gas density. 
In a comparison against observational constraints of dark matter halo
triaxiality for the Milky Way we find that $20\%$ of the galaxies in
our sample are consitent with observational results that favor
$b/a\approx 1$ and $c/a>0.77$ but are at odds with other results that
favor $c/a\approx0.44$. 
\end{abstract}

>>>>

\title[The shape of dark matter haloes at $z=0$ in Auriga]
{Dark matter halo shape at $z=0$ in the Auriga simulations: radial
  evolution and alignment with the stellar disk}
\author[Jesus Prada,  Jaime E. Forero-Romero, Volker Springel ]{
Jesus Prada,$^{1}$\thanks{E-mail: jd.prada1760@uniandes.edu.co}
Jaime E. Forero-Romero,$^{1}$
Volker Springel$^{2}$
\\
% List of institutions
$^{1}$Departamento de F\'sica, Universidad de los Andes, Cra. 1 No.
18A-10, Edificio Ip, Bogot\'a, Colombia.\\
$^{2}$Heidelberg Institute for Theoretical Studies,
Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany.\\
}

@@

% Enter the current year, for the copyright statements etc.
\pubyear{2019}

@@

% Abstract of the paper
\begin{abstract}
We present shape measurements of dark matter halos at redshift $z=0$
in a suite of 30 cosmological simulations from the Auriga project. 
We compare the results in full magnetohydrodynamics against dark
matter only physics. 
We find a strong influence of baryons in making dark matter halos
rounder at all radii compared to its dark matter only counterparts.
This change in triaxility is more pronounced towards the inner
regions of the halo.
In simulations with baryons we measure a twist in the dark matter
shape as a function of radial distance; this effect is almost absent
in dark matter only simulations.
We quantify this twisting effect as a changing degree of alignment
bewteen the halo shape with the stellar disk shape as a function of radius.
An almost perfect alignment between the two components appears at
$0.25R_{200}\sim56$ kpc.  
We show that the twist in the radial shape evolution correlates
positively with triaxility.
Furthermore, at distances $\lesssim 30$ kpc, rounder and coherent dark
matter distributions correlate with extended massive stellar disks and
low gas densities. 
In a comparison against observational constraints of dark matter halo
triaxiality for the Milky Way we find that $20\%$ of halos our sample
are consistent with observational results derived from the Pal 5 stream
that favor $b/a\approx 1$ and $c/a>0.77$. Including baryons is a
required element to achieve this level of agreement.
In contrast, all simulations (dark matter only and with baryons) are at
odds with the constrains derived from the Saggitarius stream that
favor $b/a\approx 0.97$ and $c/a\approx0.44$.
\end{abstract}
